Exploration

### 2018-03-05 exploration.Rmd
added theoretical immigration

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
mean_abund <- 2000
regPlots <- F
ipmPlots <- F
z.rng <- c(1, 12)
n <- 100  # ncells in IPM matrix
cells <- 1:n_distinct(lc.df$CellID)
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
                    s_x=c(1, -.1, -.5, -.1, .05, -2, -.4, -.4), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(1, -.1, 1.5, -.1, .1, -2), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
                    fl_x=c(1, -.01, -1, -.01, .2, .2), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, 1, -.1, .2), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.02, # p(establishment)
                    rcr_SB=0.2, # p(recruit from seedbank)
                    rcr_dir=0.6, # p(recruit directly)
                    s_SB=0.3, # p(survive in seedbank additional year)
                    p_emig=0.3, # p(seed emigrates)
                    sdd_max=4, # max SDD distance in cells
                    sdd_rate=0.8 # SDD dispersal rate
                    ))
```

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
                                 y=1:max(lc.df$y))) %>%
  mutate(x_y=paste(x, y, sep="_")) %>%
  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
         temp2=temp^2,
         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
         prec2=prec^2,
         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
         inbd=!is.na(match(.$x_y, lc.df$x_y)),
         id=row_number(), 
         id.inbd=min_rank(na_if(inbd*id, 0)),
         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
env.rct[is.na(env.rct)] <- 0
sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
                        g.p=list(sdd.max=sp_i$p$sdd_max, 
                                 sdd.rate=sp_i$p$sdd_rate, 
                                 bird.hab=rep(1,5)))
env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
rm(env.rct)
## sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,546])) + geom_tile()
```

@@

  #---- Truth: use assigned slopes to fill IPM matrix
  ### growth and survival
  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_gz, X.g=X_g[i,])
  S.v <- calc_surv(y, p=p, n_sz=n_sz, X.s=X_s[i,])
  ### Fix eviction of individuals below/above size limits
  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
  ### fill P and F matrices
  for(k in z.i) Ps[k,z.i,i] <- G.mx[k-1,]*S.v
  Ps[1,1,i] <- calc_staySB(p)
  Fs[z.i,z.i,i] <- h*outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz,
                             n_flz=n_flz, X.seed=X_seed[i,], X.fl=X_fl[i,])
  Fs[1,z.i,i] <- calc_addSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                            X.seed=X_seed[i,], X.fl=X_fl[i,])
  Fs[z.i,1,i] <- calc_rcrSB(y, p)

  #---- generate data
  ## survival
  E_surv <- antilogit(size.mx[,1:n_sz] %*% p$s_z + c(X_s[i,] %*% p$s_x))
  d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)
  ## growth
  E_grow <- size.mx[,1:n_gz] %*% p$g_z + c(X_g[i,] %*% p$g_x)
  d[[i]]$sizeNext <- rnorm(n_i[i], E_grow, p$g_sig)
  d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)
  ## flowering 
  E_fl <- antilogit(size.mx[,1:n_flz] %*% p$fl_z + c(X_fl[i,] %*% p$fl_x))
  d[[i]]$fl <- rbinom(n_i[i], 1, E_fl)
  ## seed production
  E_seed <- d[[i]]$fl * 
    exp(size.mx[,1:n_seedz] %*% p$seed_z + c(X_seed[i,] %*% p$seed_x))
  d[[i]]$seed <- rpois(n_i[i], E_seed)
  d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
  d[[i]]$seed <- ifelse(d[[i]]$fl==1, d[[i]]$seed, NA)
  nSd[i] <- sum(d[[i]]$seed, na.rm=T)
  ## seed bank: native
  B_1[i] <- B[i] * p$s_SB * (1 - p$rcr_SB) + 
    nSd[i] * (1 - p$p_emig) * (1 - p$rcr_dir) * p$s_SB
  ## recruits: native
  rcr.n <- max(ceiling(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p$p_est +
                   B[i] * p$rcr_SB * p$p_est), 2) # ≥ 2 for IPMpack sd(rcr_z')
  rcr.i <- (1:rcr.n) + n_i[i]
  d[[i]][rcr.i,] <- NA
  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.n, p$rcr_z[1], p$rcr_z[2])
  ## emigrants
  sdd[,,3,i] <- nSd[i] * p$p_emig * sdd[,,1,i]
}
for(i in 1:n.cell) {
  id.i <- env.df$id[i]

  #---- Truth: add immigrants

  #---- generate data: add immigrants
  # total immigrants
  D[i] <- sum(sdd[,,3,][sdd[,,2,]==id.i])
  # seed bank: immigrants
  B_1[i] <- B_1[i] + D[i] * (1 - p$rcr_dir) * p$s_SB
  # recruits: immigrants
  rcr.imm <- max(ceiling(D[i] * p$rcr_dir * p$p_est), 2)
  rcr.i <- (1:rcr.imm) + nrow(d[[i]])
  d[[i]][rcr.i,] <- NA
  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.imm, p$rcr_z[1], p$rcr_z[2])
}
IPMs <- Ps + Fs
lam.df <- env.df
lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% add_column(nSd=nSd, B0=B, B1=B_1, D=D)

>>>>

## Species description: Life history, characteristics, relevant attributes
The simulated species is fully described here. The object `sp_i` is a list containing elements for the global parameters, relationships with environmental variables, and demographic matrices that are applicable or potentially applicable for that modeling strategy. Parameters are abbreviated with the associated kernel component (e.g., `s_*` for survival, `g_*` for growth, etc) and the type of covariate (`*_z` for size, `*_x` for environmental covariates).
```{r sp_i}
mean_abund <- 2000
regPlots <- F
ipmPlots <- F
z.rng <- c(1, 12)
n <- 100  # ncells in IPM matrix
cells <- 1:n_distinct(lc.df$CellID)
n.cell <- length(cells)
n_i <- rpois(n.cell, mean_abund)
sp_i <- list(p=list(s_z=c(-1.75, .75), # b1 + b2*z
                    s_x=c(1, -.1, -.5, -.1, .05, -2, -.4, -.4), # b1*x1 + ...
                    g_z=c(.2, 2, -0.1), # b1 + b2*z + b3*z^2
                    g_x=c(1, -.1, 1.5, -.1, .1, -2), #b1*x1 + ...
                    g_sig=1, # growth ~ N(E, g_sig)
                    fl_z=c(-1.5, .1, .1), # b1 + b2*z + b3*z^2 + b4*z^3
                    fl_x=c(1, -.01, -1, -.01, .2, .2), # b1*x1 + ...
                    seed_z=c(2, 0.5, -.03), # b1 + b2*z + b2*z^2
                    seed_x=c(1, -.1, 1, -.1, .2), # b1*x1 + ...
                    rcr_z=c(1.5, 0.4), # N(mean=rcrt1, sd=rcrt2)
                    p_est=0.02, # p(establishment)
                    rcr_SB=0.2, # p(recruit from seedbank)
                    rcr_dir=0.6, # p(recruit directly)
                    s_SB=0.3, # p(survive in seedbank additional year)
                    p_emig=0.999, # p(seed emigrates)
                    sdd_max=3, # max SDD distance in cells
                    sdd_rate=0.8 # SDD dispersal rate
                    ))
```

## Landscape
The landscape characteristics are aggregated here. It is a gridded landscape with climate and LULC data.
```{r landscape}
env.rct <- as.tibble(expand.grid(x=1:max(lc.df$x),
                                 y=1:max(lc.df$y))) %>%
  mutate(x_y=paste(x, y, sep="_")) %>%
  mutate(temp=c(scale(lc.df$bio1_mean))[match(.$x_y, lc.df$x_y)],
         temp2=temp^2,
         prec=c(scale(lc.df$bio12_mean))[match(.$x_y, lc.df$x_y)],
         prec2=prec^2,
         pOpn=lc.df$nlcd1_mean[match(.$x_y, lc.df$x_y)],
         pOth=lc.df$nlcd2_mean[match(.$x_y, lc.df$x_y)],
         pDec=lc.df$nlcd3_mean[match(.$x_y, lc.df$x_y)],
         pEvg=lc.df$nlcd4_mean[match(.$x_y, lc.df$x_y)],
         pMxd=lc.df$nlcd5_mean[match(.$x_y, lc.df$x_y)],
         inbd=!is.na(match(.$x_y, lc.df$x_y)),
         id=row_number(), 
         id.inbd=min_rank(na_if(inbd*id, 0)),
         lat=lc.df$top[match(.$x_y, lc.df$x_y)],
         lon=lc.df$left[match(.$x_y, lc.df$x_y)])
env.rct[is.na(env.rct)] <- 0
sdd.pr <- sdd_set_probs(ncell=n.cell, lc.df=env.rct, lc.col=8:12,
                        g.p=list(sdd.max=sp_i$p$sdd_max, 
                                 sdd.rate=sp_i$p$sdd_rate, 
                                 bird.hab=rep(1,5)))
env.df <- filter(env.rct, inbd) %>% select(4:12, 1:3, 13:17) %>%
  mutate(pOpn=c(scale(pOpn)), pOth=c(scale(pOth)), pDec=c(scale(pDec)),
         pEvg=c(scale(pEvg)), pMxd=c(scale(pMxd)))
rm(env.rct)
## sdd.pr[,,2,] has indexes based on env.df$id -- NOT env.df$id.inbd
# ggplot(env.df, aes(x=lon, y=lat, fill=id %in% sdd.pr[,,2,760])) + geom_tile()
```

@@

  #---- Truth: use assigned slopes to fill IPM matrix
  ### growth and survival
  G.mx <- h*outer(y, y, calc_grow, p=p, n_gz=n_gz, X.g=X_g[i,])
  S.v <- calc_surv(y, p=p, n_sz=n_sz, X.s=X_s[i,])
  ### Fix eviction of individuals below/above size limits
  for(k in 1:(n/2)) G.mx[1,k] <- G.mx[1,k] + 1 - sum(G.mx[,k])
  for(k in (n/2+1):n) G.mx[n,k] <- G.mx[n,k] + 1 - sum(G.mx[,k])
  ### fill P and F matrices
  for(k in z.i) Ps[k,z.i,i] <- G.mx[k-1,]*S.v
  Ps[1,1,i] <- calc_staySB(p)
  Fs[z.i,z.i,i] <- h*outer(y, y, calc_rcrDir, p=p, n_seedz=n_seedz,
                             n_flz=n_flz, X.seed=X_seed[i,], X.fl=X_fl[i,])
  Fs[1,z.i,i] <- calc_DirSB(y, p=p, n_seedz=n_seedz, n_flz=n_flz, 
                            X.seed=X_seed[i,], X.fl=X_fl[i,])
  Fs[z.i,1,i] <- calc_rcrSB(y, p)

  #---- generate data
  ## survival
  E_surv <- antilogit(size.mx[,1:n_sz] %*% p$s_z + c(X_s[i,] %*% p$s_x))
  d[[i]]$surv <- rbinom(n_i[i], 1, E_surv)
  ## growth
  E_grow <- size.mx[,1:n_gz] %*% p$g_z + c(X_g[i,] %*% p$g_x)
  d[[i]]$sizeNext <- rnorm(n_i[i], E_grow, p$g_sig)
  d[[i]]$sizeNext <- ifelse(d[[i]]$surv, d[[i]]$sizeNext, NA)
  ## flowering 
  E_fl <- antilogit(size.mx[,1:n_flz] %*% p$fl_z + c(X_fl[i,] %*% p$fl_x))
  d[[i]]$fl <- rbinom(n_i[i], 1, E_fl)
  ## seed production
  E_seed <- d[[i]]$fl * 
    exp(size.mx[,1:n_seedz] %*% p$seed_z + c(X_seed[i,] %*% p$seed_x))
  d[[i]]$seed <- rpois(n_i[i], E_seed)
  # d[[i]]$fl <- ifelse(d[[i]]$surv, d[[i]]$fl, NA)
  d[[i]]$seed <- ifelse(d[[i]]$fl==1, d[[i]]$seed, NA)
  nSd[i] <- sum(d[[i]]$seed, na.rm=T)
  ## seed bank: native
  B_1[i] <- B[i] * p$s_SB * (1 - p$rcr_SB) + 
    nSd[i] * (1 - p$p_emig) * (1 - p$rcr_dir) * p$s_SB
  ## recruits: native
  rcr.n <- max(ceiling(nSd[i] * (1 - p$p_emig) * p$rcr_dir * p$p_est +
                   B[i] * p$rcr_SB * p$p_est), 2) # ≥ 2 for IPMpack sd(rcr_z')
  rcr.i <- (1:rcr.n) + n_i[i]
  d[[i]][rcr.i,] <- NA
  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.n, p$rcr_z[1], p$rcr_z[2])
  ## emigrants
  sdd[,,3,i] <- nSd[i] * p$p_emig * sdd[,,1,i]
}
for(i in 1:n.cell) {
  sdd.i <- env.df$id[i] # full rectangle grid id for SDD indexes

  #---- Truth: add immigrants
  n.sdd <- nrow(which(sdd[,,2,]==sdd.i, arr.ind=TRUE))
  sdd.j <- which(sdd[,,2,]==sdd.i, arr.ind=TRUE)[,3]
  p.ij <- sdd[,,1,sdd.j][sdd[,,2,sdd.j]==sdd.i]
  for(j in 1:n.sdd) {
    Fs[z.i,z.i,i] <- Fs[z.i, z.i,i] + 
      h*outer(y, y, calc_rcrImm, p=p, p.ij[j], n_seedz=n_seedz, n_flz=n_flz,
              X.seed=X_seed[sdd.j[j],], X.fl=X_fl[sdd.j[j],])
    Fs[1,z.i,i] <- Fs[1,z.i,i] + 
      calc_ImmSB(y, p=p, p.ij[j], n_seedz=n_seedz, n_flz=n_flz,
                            X.seed=X_seed[sdd.j[j],], X.fl=X_fl[sdd.j[j],])
  }

  #---- generate data: add immigrants
  # total immigrants
  D[i] <- sum(sdd[,,3,][sdd[,,2,]==sdd.i])
  # seed bank: immigrants
  B_1[i] <- B_1[i] + D[i] * (1 - p$rcr_dir) * p$s_SB
  # recruits: immigrants
  rcr.imm <- max(ceiling(D[i] * p$rcr_dir * p$p_est), 2)
  rcr.i <- (1:rcr.imm) + nrow(d[[i]])
  d[[i]][rcr.i,] <- NA
  d[[i]]$sizeNext[rcr.i] <- rnorm(rcr.imm, p$rcr_z[1], p$rcr_z[2])
}
IPMs <- Ps + Fs
lam.df <- env.df
lam.df$lambda <- apply(IPMs, 3, function(x) Re(eigen(x)$values[1]))
lam.df %<>% add_column(nSd=nSd, B0=B, B1=B_1, D=D)
